---
title: "Outputting Figures"
description: >
  Outputting Figures.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Outputting Figures}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
resource_files:
  - pdf/MyFigure.pdf
  - pdf/MyRecordedFigure.pdf
---


## Introduction

This vignette demonstrates how to output figures to RTF using the
`fig_to_rtf` function. The `fig_to_rtf` function takes a plot object as input and 
outputs an RTF, the format of which is fairly opinionated, as you will see.

## Get started

First load the package and format some example data.

```{r setup, message=FALSE, warning=FALSE, eval=FALSE}
library(ggplot2)
library(r2rtfExtras)

# Create ggplot object
plt <- ggplot(mtcars, aes(x=mpg, y=disp)) +
  geom_point() +
  theme_classic()

```

Next, we'll use the `fig_to_rtf` function with defaults, providing only the
minimally required arguments and figure size arguments.

```{r example1, message=FALSE, warning=FALSE, eval=FALSE}

fig_to_rtf(plot_obj        = plt,
           file_name       = "MyFigure.rtf", 
           figure_title    = "Figure 1. My figure",
           footer_snapshot = "Analysis 2025",
           path_program    = "C:\\\\program\\\\path\\\\code.r",
           fig_width       = 7, 
           fig_height      = 3.2)
```


```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/MyFigure.pdf")
```


It is also possible to output figures created using base R plotting capabilities.
You can record the plot using `recordPlot()`, a function from the `grDevices` packages,
set the class, and output. An example is shown below.


```{r example2, message=FALSE, warning=FALSE, eval=FALSE}

plot(mtcars$mpg, mtcars$disp, pch=16)
plt_record <- recordPlot()
class(plt_record) <- c(class(plt_record), "ggplot")

fig_to_rtf(plot_obj        = plt_record,
           file_name       = "MyRecordedFigure.rtf", 
           figure_title    = "Figure 1. My recorded figure",
           footer_snapshot = "Analysis 2025",
           path_program    = "C:\\\\program\\\\path\\\\code.r",
           fig_width       = 7, 
           fig_height      = 5)
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/MyRecordedFigure.pdf")
```


For more information on inputs, check out the function documentation `?fig_to_rtf`.

