---
title: "Outputting Tables"
description: >
  Outputting Tables.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Outputting Tables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
resource_files:
  - pdf/MyTable.pdf
  - pdf/MyTableImproved.pdf
  - pdf/MyColSpanTable.pdf
  - pdf/MyColSpanTableProblem.pdf
---


## Introduction

This vignette demonstrates how to output dataframes to RTF using the
`df_to_rtf` function. The `df_to_rtf` function takes a dataframe as input and 
outputs an RTF, the format of which is fairly opinionated, as you will see.

## Get started

First load the package and format some example data.

```{r setup, message=FALSE, warning=FALSE, eval=FALSE}
library(r2rtfExtras)

df       <- mtcars[1:10,]
df$Model <- row.names(df)
df       <- df[c("Model", "mpg", "cyl", "disp", "hp")]
```

Next, we'll use the `df_to_rtf` function with defaults, providing only the
minimally required arguments.

```{r example1, message=FALSE, warning=FALSE, eval=FALSE}

df_to_rtf(df              = df,
          file_name       = "MyTable.rtf",
          table_title     = "Table 1. MTCARS Dataset",
          footer_snapshot = "Henderson and Velleman (1981)", 
          path_program    = "C:\\\\program\\\\path\\\\tables.Rmd")
```


```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/MyTable.pdf")
```

There are some enforced opinions in the output: the page header has a right-aligned
page number and the page footer has a left-aligned snapshot and a right-aligned timestamp,
as well as rows for the paths to the program and output.

You should also notice in the output:

- The page header is largely empty as it only includes the page number.
- The table column names are not very informative.
- The page footer Output Name is less than desirable.

Let's improve on these issues to generate a more informative RTF:

```{r example2, message=FALSE, warning=FALSE, eval=FALSE}

df_to_rtf(df              = df,
          file_name       = "MyTableImproved.rtf",
          org             = "My Organization",
          header_page     = "My MTCARS Analysis",
          table_title     = "Table 1. MTCARS Dataset",
          header_table1   = c("Model", "Miles Per Gallon", "Cylinders", 
                              "Displacement", "Horsepower"), 
          footer_snapshot = "Henderson and Velleman (1981)", 
          path_program    = "C:\\\\program\\\\path\\\\tables.Rmd",
          path_output     = "C:\\\\output\\\\path\\\\MyTableImproved.rtf")
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/MyTableImproved.pdf")
```

Much better!

Finally, a few things that are not readily apparent. By default, the 
`df_to_rtf` function will add the following:

- A stylesheet that allows for setting the table title to a heading style of level 1 through 4.
- A table title that is set to the heading style of level 2 by default and is recognized by Word.
- Bookmark start and end tags around the table. Very useful for auto-linking within other documents.


## Tables with column spanning headers

We can also output tables with column spanning headers. Let's see how this works:

```{r example3, message=FALSE, warning=FALSE, eval=FALSE}

df_to_rtf(df              = df,
          file_name       = "MyColSpanTable.rtf",
          org             = "My Organization",
          header_page     = "My MTCARS Analysis",
          table_title     = "Table 1. MTCARS Dataset",
          header_table0   = c(" ", "Performance", "Engine"),
          header_table1   = c("Model", "Miles Per Gallon", "Cylinders", 
                              "Displacement", "Horsepower"), 
          col_widths0     = c(1, 4, 4),
          col_widths1     = c(1, 2, 2, 2, 2),
          col_just0       = c("l", "c", "c"),
          col_just1       = c("l", "c", "c", "c", "c"),
          footer_snapshot = "Henderson and Velleman (1981)", 
          path_program    = "C:\\\\program\\\\path\\\\tables.Rmd",
          path_output     = "C:\\\\output\\\\path\\\\MyColSpanTable.rtf")
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/MyColSpanTable.pdf")
```

Currently, `df_to_rtf` only allows for one level of column spanning, as shown in
the example output above. Here are some key details:

- `header_table0` specifies the column spanning header row.
- `header_table1` specifies the column names for the second header row, which is the row that is directly above the data rows.
- `col_widths0` and `col_widths1` must sum to the same value. The alignment between the spanning row and the column names is largely controlled by the `col_widths0` and `col_widths1` arguments.
- `col_just0` is the justification of the column spanning header row.
- `col_just1` is the justification of the column names row and the data rows.

What happens if the inputs are not quite consistent? Below, the `col_widths0` 
and `col_widths1` sum to the same value, but the header row is not consistent
with the spanning row.

```{r example4, message=FALSE, warning=FALSE, eval=FALSE}

df_to_rtf(df              = df,
          file_name       = "MyColSpanTableProblem.rtf",
          org             = "My Organization",
          header_page     = "My MTCARS Analysis",
          table_title     = "Table 1. MTCARS Dataset",
          header_table0   = c(" ", "Performance", "Engine"),
          header_table1   = c("Model", "Miles Per Gallon", "Cylinders", 
                              "Displacement", "Horsepower"), 
          col_widths0     = c(1, 4, 4),
          col_widths1     = c(1, 2.2, 1.7, 2.1, 2),
          col_just0       = c("l", "c", "c"),
          col_just1       = c("l", "c", "c", "c", "c"),
          footer_snapshot = "Henderson and Velleman (1981)", 
          path_program    = "C:\\\\program\\\\path\\\\tables.Rmd",
          path_output     = "C:\\\\output\\\\path\\\\MyColSpanTableProblem.rtf")
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/MyColSpanTableProblem.pdf")
```

As you can see above, the spanning header row, Performance, has a right border
that is sligthly misaligned due to inconsistency between the `col_widths0` and 
`col_widths1` arguments.

For more information on inputs, check out the function documentation `?df_to_rtf`.

